Viscous Erosion in Porous Media

### 2018-03-26 docs/methods/erosion_methods.tex
edits to paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Contributions} The present study develops numerical methods
to simulate erosion of multiply-connected two-dimensional geometries.
In this fashion, individual pores are resolved rather than
homogenized.  A boundary integral equation (BIE) is used to solve the
incompressible Stokes equations in the complex geometry.  We resolve
individual grains and the non-negligible interactions between bodies.
To link the eroding body and fluid dynamics, the body is eroded in the
normal direction at a rate that is proportional to the magnitude of the
shear stress.

To create high-fidelity and computationally efficient numerical methods,
two fast and accurate methods are employed.  First, the fluid velocity
and the erosion rate (shear stress) are efficiently computed with
spectral accuracy by using a fast-multipole-accelerated BIE method.  The
required spatial derivatives are computed with spectral accuracy in
Fourier space.  In addition to compute the shear stress, we also compute
the vorticity in the fluid bulk.  The vorticity is
particularly insightful since it is identical to the shear stress on
no-slip boundaries.  Therefore, by computing the vorticity, we are able
to locate regions where the erosion rates will be fastest or slowest
(Fig.~\ref{fig:50bodies}).

Once the erosion rate is computed, the shape is integrated in time with
second-order accuracy.  Tracking the interface of fluid-structure
interaction problems with a long time horizon is particularly
difficult---the mesh is often distorted and tangled by the dynamics and
regions of high curvature develop.  These two issues are resolved by,
first, introducing a tangential velocity field to maintain a grid that
is equispaced in arclength, but that does not affect the dynamics, and,
second, a curvature penalization term.  Furthermore, the frame of
reference is changed to \thL~variables so that the curvature term
appears as a diffusive term that can easily be handled in a numerically
stable fashion, e.g.~via an exponential integrator or an implicit
method.

We validate our method on a single circular body placed symmetrically in
a pipe flow.  By artificially fixing the area of the body as a function
of time, we are able to characterize the long-time behavior of the shape
and the resulting flow and these results agree with analytical results.
Then, we compute the vorticity of erosion in multiply-connected
geometries to better characterize erosion in porous media.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Contributions} The present study develops numerical methods
to simulate erosion of multiply-connected two-dimensional geometries.
In this fashion, individual pores are resolved rather than
homogenized.  A boundary integral equation (BIE) is used to solve the
incompressible Stokes equations in the complex geometry.  The resulting
method resolves individual grains and the non-negligible interactions
between bodies.  To link the eroding body and fluid dynamics, the body
is eroded in the normal direction at a rate that is proportional to the
magnitude of the shear stress.

To create high-fidelity and computationally efficient numerical methods,
two fast and accurate methods are employed.  First, the fluid velocity
and the erosion rate (shear stress) are computed with a
fast-multipole-accelerated Kryolv method applied to a spectrally
accurate discretization of the governing equations.  In particular,
GMRES~\cite{saa-sch1986} is applied to a spectrally-accurate Nystr\"om
discretization of the BIE~\cite{tre-wei2014}, a spectrally accurate
quadrature rule is used to compute the shear stress~\cite{sid-isr1988},
and all arclength derivatives are computed with Fourier differentiation.
In addition to compute the shear stress, we also compute the vorticity
in the fluid bulk.  The vorticity is particularly insightful since it is
identical to the shear stress on no-slip boundaries.  Therefore, by
computing the vorticity, we are able to identify regions where the
erosion rate will be fast or slow (Fig.~\ref{fig:50bodies}).

Once the erosion rate is computed, the shape is integrated in time with
second-order accuracy.  Tracking the interface of fluid-structure
interaction problems with a long time horizon is particularly
difficult---the mesh is often distorted and tangled.  Moreover, erosion
naturally leads to regions of high or infinite (i.e.~corners) curvature.
These interface tracking issues are resolved by, first, introducing a
tangential velocity field to maintain a grid that is equispaced in
arclength, but that does not affect the dynamics, and, second, a tunable
term that penalizes regions of high curvature.  Furthermore, the frame
of reference is changed to \thL~variables so that the curvature term
appears as a diffusive term that can easily be handled in a numerically
stable fashion with an exponential integrator.

We validate our method on a single circular body placed symmetrically in
a pipe flow.  By artificially fixing the area of the body as a function
of time, we are able to characterize the long-time behavior of the shape
and the resulting flow and these results agree with analytical results.
Then, we compute the vorticity of erosion in multiply-connected
geometries to better characterize erosion in porous media. 
\todo[inline]{Bryan is here in editing}
